Child Signals of Need in Utila, Honduras: An Exploratory Study

### 2024-09-09 Paper.qmd
Paper edits and addition of code to plot effects of crying on weight and height gain.

A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. This hypothisized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this (cite). However, our small sample size with signaling data and anthropometric data at two time periods limits our ability to do this in this study. For this reason, all analyses for this study were exploratory and meant to inform future studies of child signaling.

@@

: Table showing variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day. {#tbl-freqvars}

@@

: Table showing conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include SadFreqN, CryFreqN, TantrumFreqN, RunawayFreqN, ConflictFreqN, and AlloparentingFreqN. {#tbl-convertedfreqs}

@@

: Table showing variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. {#tbl-relvars}

@@

### Perceptions of child immune function (or current susceptibility to illness)

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create the CurrentHealthMean variable. In the `r sum((!is.na(modeldf$Currenthealth2)) & is.na(modeldf$Currenthealth3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2. See Table @tbl-healthvars for the wording of these variables.

|                |                                                                                       |
|--------------------------|----------------------------------------------|
| Variable       | Survey question wording                                                               |
| Currenthealth1 | If an illness is 'going around', my child will get it.                                |
| Currenthealth2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| Currenthealth3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

: Table showing variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for Currenthealth1 and Currenthealth2 were inverted so that 5 corresponded with the lowest susceptibility to illness. {#tbl-healthvars}

@@

Triceps and subscapular skinfold thicknesses were measured using skinfold calipers. Measurements were taken directly on the skin. The triceps measurement occurred at the midpoint of the upper arm. The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna]. Add flexed bicep measurementxxxxxx.

@@

For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011]. The resulting variables which represent how each child differed from the average child for their given sex and age.

Summed body fat

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables while being cognizant of the risk of overfitting the data. The payoff to this is the identification of theoretically relevant effects that should be tested in the future.

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables involved over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with household as a random effect. Control variables differed in some models, but always include child age and sex as controls. Specific control variables are mentioned in the results.

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for household. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS with household as a random effect.

@@

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  Currenthealth1 = "Currenthealth1",
  Currenthealth2 = "Currenthealth2",
  Currenthealth3 = "Currenthealth3",
  CurrentHealthMean = "Mean of current health measures",
  HeightZ = "Height Z-score (CDC)",
  WeightZ = "Weight Z-score (CDC)",
  BMIZ = "BMI Z-score (CDC)",
  SubscapMean = "Subscapular skinfold thickness (mm)",
  SubscapR = "Subscapular skinfold residuals",
  TricepMean = "Tricep skinfold thickness (mm)",
  TricepR = "Tricep skinfold residuals",
  BodyFat =  "Body fat composite"
)

@@

Child sadness, crying, and temper tantrums were reasonably common. Caregivers reported that (`r sum(na.omit(modeldf$SadFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\]) children were sad at least once a week, (`r sum(na.omit(modeldf$CryFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\]) cried at least more than once a week, and (`r sum(na.omit(modeldf$TantrumFreqN >= 8))` \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..

@@

For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.

@@

```{r}
#| label: tbl-main_models
#| tbl-cap: Regression tables for our main exploratory models.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare"
)
modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.

@@

There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child immune function, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.

@@

```{r}
#| label: tbl-conflict_dep_var
#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  UnwantedTask = "Unwanted task",
  Punishment21 = "Punishment",
  Family21 = "Family conflict",
  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
  CaregiverResponse = "Caregiver Response",
  OlderKids = "Older children",
  YoungerKids = "Younger children",
  AdultsNoChildcare = "Adults that do not provide childcare",
  AdultsChildcare = "Adults that provide childcare",
  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
)
modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

@@

For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child immune function. The slope of these relationships were positive for girls and flat or slightly negative for boys. In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys. See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.

@@

```{r}
#| label: tbl-alloparenting_dep_var
#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  UnwantedTask = "Unwanted task",
  Punishment21 = "Punishment",
  Family21 = "Family conflict",
  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
  CaregiverResponse = "Caregiver Response",
  OlderKids = "Older children",
  YoungerKids = "Younger children",
  AdultsNoChildcare = "Adults that do not provide childcare",
  AdultsChildcare = "Adults that provide childcare",
  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
)
modelsummary(malloparenting, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child immune function, an alloparenting effort by sex interaction and a body fat by sex interaction. See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots. Subscapular and tricep skinfold thicknesses were also significant predictors of summed signal cost and tantrum frequency when included in the same models as the sole measure of body fat. See xxxxSI TABLESxxxx for regression tables.

@@

```{r}
#| label: tbl-bodyfat_models
#| tbl-cap: Regression tables for the models examining the effect of body fat on signaling.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  CurrentHealthMean = "Percieved immunity",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)"
)
modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

@@

>>>>

A major challenge when studying signals of need is the fact that such signaling is expected to 1) be more common in times when individuals can benefit from increased investment (e.g., when they are facing major adversity or when a fitness-relevant opportunity arises) and 2) result in benefits to the signaler on average. This hypothisized causal, bi-directional relationship between investment and signaling makes it difficult to isolate the effects of signaling behavior on investment in children when using standard measures of child welfare (i.e., BMI, health metrics, and height for age). When experiments are not appropriate, longitudinal studies with carefully selected variables can help address this [@arjas2004]. However, our small sample size with signaling data and anthropometric data at two time periods limits our ability to do this in this study. For this reason, all analyses for this study were exploratory and meant to inform future studies of child signaling.

@@

: Variable names and question wording for measures about the frequency of behaviors. All questions had the following labels attached to each point on its Likert scale: Never; Once a month or less; More than once a month but less than once a week; More than once a week but not daily; Daily; Multiple times per day. {#tbl-freqvars}

@@

: The conversions from Likert scale to count value for the frequencies of child behaviors included in this study. Variables created with this scheme include the frequencies of sadness, crying, temper tantrums, conflict with primary caretakers, and child alloparenting effort.

@@

: Variable names and question wording for measures about relative need and investment. Both questions had the following labels attached to each point on its Likert scale: Much less; Less; The same amount; More; Much more. {#tbl-relvars}

@@

### Perceptions of child susceptibility to illness (or current susceptibility to illness)

Caretakers were also asked three questions using a five point Likert scale about their child's susceptibility to illness. We inverted the scores on two measures so that a score of 5 corresponded with the lowest susceptibility to illness before we averaged the scores of these to create a mean illness susceptibility variable. In the `r sum((!is.na(modeldf$IllnessSusceptibility2)) & is.na(modeldf$IllnessSusceptibility3))` instance in which we only data for 2 of the measures, we still calculated the mean but used a numerator of 2. See Table @tbl-healthvars for the wording of these variables.

|                        |                                                                                       |
|--------------------------|----------------------------------------------|
| Variable               | Survey question wording                                                               |
| IllnessSusceptibility1 | If an illness is 'going around', my child will get it.                                |
| IllnessSusceptibility2 | In general, my child is very susceptible to colds, flu, and other infectious diseases |
| IllnessSusceptibility3 | My child’s immune system protects him/her from most illnesses that other kids get.    |

: Variable names and question wording for measures about a child's current likelihood of becoming sick. All three questions were asked using a Likert scale from 1-5 where 1 was labeled "Strongly disagree" and 5 was labeled "Strongly agree." The scores for IllnessSusceptibility1 and IllnessSusceptibility2 were inverted so that 5 corresponded with the lowest susceptibility to illness. {#tbl-healthvars}

@@

Triceps and subscapular skinfold thicknesses were measured using skinfold calipers. Measurements were taken directly on the skin. The triceps measurement occurred at the midpoint of the upper arm. The subscapular measurement occured immediately below the inferior angle of the scapula, being picked up diagonally at the natural fold line of the skin [@antóna]. Flexed bicep circumference was measured perpendicular to the long axis of the upper arm. For this measure participants were asked to sit/stand? an position their upper arm perpendicular to their body with their elbow being help at a 90° angle, before flexing their clenched fist towards themselves as much as they could.

@@

For our independent variables of skinfold thickness, flexed bicep circumference, and grip strength on signaling behavior, we used the residuals from generalized additive models which modeled the effect of sex and age on each measurement using the mgcv package [@wood2011]. The resulting variables which represent how each child differed from the average child for their given sex and age. We also centered and scaled the skinfold thickness residuals before summing them to create an overall body fat variable.

@@

All statistical analyses were exploratory and involved fitting many models with different combinations of independent and dependent variables. While were cognizant of the risk of overfitting the data, it is still more likely than if the study was entirely confirmatory. The payoff to this approach is the identification of theoretically relevant effects that should be tested in the future.

We fit generalized linear mixed models for exploratory analyses with signal frequencies and summed signal cost as dependent variables. As our signaling variables involved over-dispersed count data, we used negative binomial regression. This was done in R with the GLMM package [@brooks2017], with household as a random effect. Control variables differed in some models, but child age and sex were always included. Specific control variables are mentioned in the results.

For the models examining 1) parental responses to the last reported time each participating child was sad or cried, 2) the relative neediness of the child, and 3) the investment a child received compared to the other children in the household, we fit ordinal logistic regression models using two methods. The first used the polr function in the MASS package [@venables2002]. The results were then interpreted with clustered standard errors for household. Due to the risk of overfitting that comes with having too few clusters [@cameron2015], we also fit the same models using BRMS [@bürkner2017] with household as a random effect.

@@

```{r sumtab_children}
#| label: tbl-sumtab_children
#| tbl-cap: Summary statistics for child study variables.
vars_c <- c(
  ChildAge = "Age (years)",
  SadFreqN = "Sad frequency (times per month)",
  CryFreqN = "Crying frequency (times per month)",
  TantrumFreqN = "Temper tantrum frequency (times per month)",
  AlloparentingFreqN = "Frequency of child's own alloparetning effort (times per month)",
  ConflictFreqN = "Frequency of conflict with focal caretaker (times per month)",
  SignalFreq = "Frequency of signals summed (times per month)",
  SignalCost = "Signal cost",
  SignalFreqMax = "Frequency of most common signal (times per month)",
  IllnessSusceptibility1 = "IllnessSusceptibility1",
  IllnessSusceptibility2 = "IllnessSusceptibility2",
  IllnessSusceptibility3 = "IllnessSusceptibility3",
  IllnessSusceptibilityMean = "Mean of current health measures",
  HeightZ = "Height Z-score (CDC)",
  WeightZ = "Weight Z-score (CDC)",
  BMIZ = "BMI Z-score (CDC)",
  SubscapMean = "Subscapular skinfold thickness (mm)",
  SubscapR = "Subscapular skinfold residuals",
  TricepMean = "Tricep skinfold thickness (mm)",
  TricepR = "Tricep skinfold residuals",
  BodyFat =  "Body fat composite"
)

@@

Child sadness, crying, and temper tantrums were common. Caregivers reported that `r sum(na.omit(modeldf$SadFreqN >= 8))` (`r signif(sum((na.omit(modeldf$SadFreqN >= 8)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%) children were sad at least once a week, `r sum(na.omit(modeldf$CryFreqN >= 8))` (`r signif(sum((na.omit(modeldf$CryFreqN >= 8)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%) cried at least more than once a week, and `r sum(na.omit(modeldf$TantrumFreqN >= 8))` (`r signif(sum((na.omit(modeldf$TantrumFreqN >= 8)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%) threw temper tantrums at least once a week. A small subset of children engaged in these signals/states at least daily (sadness = `r sum(na.omit(modeldf$SadFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$SadFreqN >= 30)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$CryFreqN >= 30)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN >= 30))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN >= 30)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). However, many children were reported to never engage in these signals/states (sadness = `r sum(na.omit(modeldf$SadFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$SadFreqN == 00)) / sum(!is.na(modeldf$SadFreqN)) * 100), 2)`%\], crying = `r sum(na.omit(modeldf$CryFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$CryFreqN == 0)) / sum(!is.na(modeldf$CryFreqN)) * 100), 2)`%\], and tantrums = `r sum(na.omit(modeldf$TantrumFreqN == 0))` children \[`r signif(sum((na.omit(modeldf$TantrumFreqN == 0)) / sum(!is.na(modeldf$TantrumFreqN)) * 100), 2)`%\]). See @fig-barplot_SignalFreq for plot showing signaling frequencies and @fig-signal_alluvial_plot for plot showing the number of individuals who engaged in specific combinations of signal frequencies..

@@

For every signaling measure except sadness (which also involves non-signaling functions), there was a statistically significant negative relationship between the age of the child and the frequency of signaling when controlling for the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics and @fig-effect_age for effects plots.

@@

```{r}
#| label: tbl-main_models
#| tbl-cap: Regression tables for our main exploratory models.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare"
)
modelsummary(list(Sadness = msadH, Crying = mcryH, Tantrum = mtantrumH, SignalFreq = msfH, SignalCost = mscH, SignalFreqMax = mmsfH), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

There were statistically significant positive relationships between the frequency of conflict between a child and their primary caretaker and all of our signaling measures when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_conflict for effects plots.

@@

There were statistically significant negative relationships between years of caregiver education and all of our signaling measures except sadness (`r stats$msadH$EducationLevelYears$str`) when controlling for the sex of the child, the age of the child, household demographics, caretaker perceptions of child susceptibility to illness, and an alloparenting effort by sex interaction. See @tbl-main_models for regression statistics @fig-effect_edu for effects plots.

@@

```{r}
#| label: tbl-conflict_dep_var
#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  UnwantedTask = "Unwanted task",
  Punishment21 = "Punishment",
  Family21 = "Family conflict",
  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
  CaregiverResponse = "Caregiver Response",
  OlderKids = "Older children",
  YoungerKids = "Younger children",
  AdultsNoChildcare = "Adults that do not provide childcare",
  AdultsChildcare = "Adults that provide childcare",
  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
)
modelsummary(mconflict, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

For each signaling measure, there was also a significant interaction between how often a child cared for their siblings and the sex of the child when controlling for the age of the child, the sex of the child, household demographics, the frequency of conflict with one's primary caretaker, and caretaker perceptions of child susceptibility to illness. The slope of these relationships were positive for girls and flat or slightly negative for boys. In other words, our models predict that as children provide more allocare within their families, the frequency of signaling would increase with alloparenting effort for girls and remain at the same rate or decrease slightly with alloparenting effort for boys. See @tbl-main_models for the corresponding regression table and @fig-effect_allo_by_sex for effects plots.

@@

@@

```{r}
#| label: tbl-alloparenting_dep_var
#| tbl-cap: Regression table for the model examining the frequency of conflict between child and caretaker as the dependent variable.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  UnwantedTask = "Unwanted task",
  Punishment21 = "Punishment",
  Family21 = "Family conflict",
  DiscomfortPainInjuryIllness = "Discomfort/Pain/Injury/Illness",
  CaregiverResponse = "Caregiver Response",
  OlderKids = "Older children",
  YoungerKids = "Younger children",
  AdultsNoChildcare = "Adults that do not provide childcare",
  AdultsChildcare = "Adults that provide childcare",
  OtherChildAlloparentingFreqN = "Other household child alloparetning freq."
)
modelsummary(malloparenting, gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

There were significant negative relationships between summed signal cost (`r stats$mBFc$BodyFat$str`) crying frequency (`r stats$mBFcry$BodyFat$str`), and tantrum frequency (`r stats$mBFtantrum$BodyFat$str`) and a scaled composite of subscapular and tricep skinfold thicknesses when controlling for the sex of the child, the age of the child, the frequency of conflict with one's primary caretaker, household demographics,caretaker perceptions of child susceptibility to illness, an alloparenting effort by sex interaction and a body fat by sex interaction. See @tbl-bodyfat_models for regression statistics and @fig-effect_bodyfat for effects plots. Subscapular and tricep skinfold thicknesses were also significant predictors of summed signal cost and tantrum frequency when included in the same models as the sole measure of body fat. See xxxxSI TABLESxxxx for regression tables.

@@

```{r}
#| label: tbl-bodyfat_models
#| tbl-cap: Regression tables for the models examining the effect of body fat on signaling.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)"
)
modelsummary(list(Sadness = mBFsad, Crying = mBFcry, Tantrum = mBFtantrum, SignalFreq = mBFf, SignalCost = mBFc, SignalFreqMax = mBFfm), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

@@

## Crying frequency is positively associated with greater height and weight gain in a small longitudinal sample

Signaling data and longitudinal data on child height and weight (2023 and 2024) exist for `r sum(table(height_long_cry$frame$HeightMean_2024))` children. In this sample, there was a significant effect of a child age by crying frequency interaction on 2024 child height and weight whereby more crying was associated with more growth when controlling for 2023 height and weight.

See @tbl-longitudinal for the corresponding regression table.

# Check: SD (Observations) in the table
```{r}
#| label: tbl-longitudinal
#| tbl-cap: Regression tables for the models examining height and weight growth.
r_vars <- c(
  ChildAge = "Age (years)",
  SexMale = "Sex (male)",
  OtherChildrenHH = "Other children",
  LogIncome = "Income (log)",
  number_adults = "Adults in home",
  PartnerStatusUnpartnered = "Partner status (single)",
  EducationLevelYears = "Parent ed. (years)",
  ConflictFreqN = "Conflict freq.",
  SignalFreq = "Signal freq.",
  SignalCost = "Signal cost",
  SignalFreqMax = "Max signal freq.",
  IllnessSusceptibilityMean = "Percieved illness susceptibility",
  SadFreqN = "Sadness freq.",
  CryFreqN = "Crying freq.",
  TantrumFreqN = "Tantrum freq.",
  AlloparentingFreqN = "Child allocare",
  TricepR = "Tricep skinfold (residuals)",
  SubscapR = "Subscapular skinfold (residuals)",
  BodyFat = "BodyFat (composite)",
  RelativeNeed3 = "Relative need (w/in household)",
  HeightMean_2024 = "2024 height",
  HeightMean_2023 = "2023 height",
  WeightMean_KG_2024 = "2024 weight",
  WeightMean_KG_2023 = "2023 weight"
)

modelsummary(list(Weight2024 = weight_long_cry, Height2024 = height_long_cry), gof_map = "nobs", coef_rename = r_vars, stars = c('*' = .05, '**' = .01, '***' = .001), coef_omit = "(Intercept)")
```

::: {#fig-effect_longitudinal_plot}
```{r}
longitudinal_plot
```

The effect of a child crying frequency by age on child weight and height in 2024, controlling for the corresponding measure in 2023. Rug corresponds to actual values of crying frequency per month with a small amount of jitter to reveal overlapped values. See @tbl-longitudinal for the corresponding regression table.
:::

@@

Signaling behavior did not predict (try more models?) differences in relative caregiver investment between children within households. However, crying frequency was  associated with both height and weight gain for older children. If older children face greater reputation costs from crying, crying may be a relatively credible credible signal compared to crying among younger children, with it being a less viable strategy to exaggerate one's value of investment or the costliness of adversity.
